Governing through patronage: the bargain for education in decentralized Brazil

### 2020-10-12 paper/paper.tex
redraft intro

Across the world, vulnerable populations rely on governments for access to basic public services such as education. In the developing world, the quality of public education remains low: functional illiteracy is widespread, and basic mathematical operations beyond the reach of many \citep{vegas_raising_2007}. The World Bank refers to this phenomenon as ``learning poverty'': stunted learning when access to a schooling exists.\footnote{For more information, see \hyperlink{here}{https://www.worldbank.org/en/topic/education/brief/learning-poverty}.}

Understanding who benefits from these services and how politicians decide on that allocation is a well-established tradition in political science \citep{boix_democracy_2003, dixit_determinants_1996}. But we have less intuition on services that are de facto open to all: public education.

Raises two related questions. The first one concerns the relationship between the quality of public services and the structure of the bureaucracy \citep{pepinsky_bureaucracy_2017, rasul2018management}. Moving up the causal ladder, the question concerns the relationship between politicians and bureaucratic institutions and how theses personnel decisions \citep{finan_personnel_2015} have downstream consequences for the quality of these services \citep{pepinsky_bureaucracy_2017}

These outcomes are the end-result of a series of interlocking steps: from personnel decisions to service provision \citep{pepinsky_bureaucracy_2017}, each link in this institutional chain of command connects political principals to the child sitting in the classroom.

>>>>

Across the world, vulnerable populations rely on governments for access to basic public services such as education. In the developing world, the quality of public education remains low: functional illiteracy is widespread, and basic mathematical operations beyond the reach of many \citep{vegas_raising_2007}. The World Bank refers to this phenomenon as ``learning poverty'': stunted learning when access to a schooling exists.\footnote{For more information, see \hyperlink{here}{https://www.worldbank.org/en/topic/education/brief/learning-poverty}.} These outcomes result from political choices, but a growing body of literature recognizes the important role bureaucratic institutions mediating these outcomes \citep{pepinsky_bureaucracy_2017,finan_personnel_2015}.

Understanding how these political decisions cascade down the bureaucracy remains a challenge. Public services the end-result of a series of interlocking steps: from personnel decisions to service provision \citep{pepinsky_bureaucracy_2017}, each link in this institutional chain of command connects political principals to the child sitting in the classroom \citep{}.
